Efficient Handling of String-Number Conversion

### 2019-11-23 draft.tex
intro - upto the beginning of the technical part (of the intro sec)

For many languages such as Java, string-number conversions are core operations, and much symbolic execution can not been done without supporting it. In fact,  a code that receives input tends to need to convert at least some of that input into numbers,  which makes the handling of string-integer conversions of the utmost importance. For example, the program fragment below is a variant of the Luhn test algorithm that is often used in credit card or personal ID validation.

@@

The input \verb|value| of the Luhn test algorithm is a sequence of digits. The algorithm processes the digits in the reversed order. The value of every odd digit (e.g., 1st, 3rd, etc.) is added to \verb|sum| directly. For the value of every even digit, the algorithm (1) doubles its value, (2) subtracts its value by $9$ if the doubled-value is larger than $9$, and (3) adds the final result to \verb|sum|. At the end, the input is validated if the string value of \verb|sum| ends by an $0$ (i.e., \verb| sum mod 10=0|).

To check whether the program path that traverses both loops exactly once, which corresponds to the case when $| \verb|value||=2$, and finally passes this test has a valid input, we create the following (string) constraint:

@@

Even, more crucially, in many programming languages, string-number conversion is a part of the definition of the core semantics of the language. 
 JavaScript, which powers most interactive content on the Web and increasingly server-side code with Node.js, is one of such languages. Other scripting languages do too, but we focus on JavaScript due to its prominence. To see how string-integer conversion pervades semantics, consider the following program:

@@

But if array indices are modeled as strings, we must handle arithmetic somehow. Let us look at the case of line 6, we need to update the value of \texttt{x["03"-1]}. The evaluation of the expression \texttt{"03"-1} involves an implicit type conversion from the string \texttt{"03"} to an integer value $3$ due to the \texttt{-} (minus) operation. The result of the evaluation of \texttt{"03"-1} is the integer $2$, which is then converted back to string \texttt{"2"} and used as the array index. Hence \texttt{x["03"-1]} means the array element of \texttt{x["2"]}. Even for a simple example like this, the conversion between string and number is unavoidable. This is a rather basic   array operation in JavaScript, and not  handling string-number conversion operations will cripple any analysis of non-trivial JavaScript code. Thus, we need stronger solvers that are  able to handle string-number conversion operations in order to be able to analyse  real code.

>>>>

%For many languages such as Java, string-number conversion is not a core operation, and much symbolic execution can be done without supporting it; however, this is not true of scripting languages, of which JavaScript is the most notable.
%For many languages such as JavaScript, string-number conversions are core operations, and symbolic execution cannot be done without supporting it.
In fact, a code that receives string input tends to need to convert at least some of that input into numbers.
%,  which makes the handling of string-integer conversions of the utmost importance.  \yfc{maybe this is a too strong argument using this example}
For example, the program fragment below is a variant of the Luhn test algorithm that is often used in credit card or ID validation.

@@

The input \verb|value| of the Luhn test algorithm is a sequence of digits. The algorithm processes the digits in the reversed order. The value of every odd digit (e.g., 1st, 3rd, etc.) is added to \verb|sum| directly. For the value of every even digit, the algorithm (1) doubles its value, (2) subtracts its value by $9$ if the doubled-value is larger than $9$, and (3) adds the final result to \verb|sum|. At the end, the input is validated if the last digit of \verb|sum| is $0$ (i.e., \verb| sum mod 10=0|).

To check whether the program path that traverses both loops exactly once and finally passes this test has a valid input, we create the following (string) constraint:

@@

Even, more crucially, in many programming languages, string-number conversion is a part of the definition of their core semantics. 
 JavaScript, which powers most interactive content on the Web and increasingly server-side code with Node.js, is one of such languages. Other scripting languages do too, but we focus on JavaScript due to its prominence. To see how string-integer conversion pervades semantics, consider the following program:

@@

But if array indices are modeled as strings, we must handle arithmetic somehow. Let us look at the case of line 6, we need to update the value of \texttt{x["03"-1]}. The evaluation of the expression \texttt{"03"-1} involves an implicit type conversion from the string \texttt{"03"} to an integer value $3$ due to the \texttt{-} (minus) operation. The result of the evaluation of \texttt{"03"-1} is the integer $2$, which is then converted back to string \texttt{"2"} and used as the array index. Hence \texttt{x["03"-1]} means the array element of \texttt{x["2"]}. Even for a simple example like this, the conversion between string and number is unavoidable. This is a rather basic   array operation in JavaScript, and not  handling string-number conversion operations will cripple any analysis of non-trivial JavaScript code. Thus, we need stronger solvers that are  able to handle string-number conversion operations in order to be able to analyze  real code.
